Manuscript_Github_Pchabaudi_pir_TPMHeatmaps

### 2021-04-21 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_pir_TPMHeatmaps.Rmd
Refined Chapl heatmap and added clustered ht

combined_data_ordered <- tibble(Geneid = rownames(combined_data_pir),
                                means = apply(combined_data_pir, 1, mean),
                                chapl_aapl_name = factor(
                                  combined_data_pir$chapl_aapl_name
                                )) %>% 
  arrange(chapl_aapl_name, 
          desc(means)) %>% 
  .$Geneid

@@

#Sub-family annotation
chapl_aapl_name_info <- combined_data_pir_ht$chapl_aapl_name
annotation_col <- c(chapl_aapl_name_col_num, 'ref' = 'black')

@@

ht <- Heatmap(
  log2(1+combined_data_pir_ht[,!colnames(combined_data_pir_ht) %in% 'chapl_aapl_name']),
  cluster_rows = FALSE,
  cluster_columns = FALSE,
  show_row_names = FALSE,
  name = 'log(TPM+1)', 
  col = colorRamp2(c(0,
                     maxcol/4,
                     maxcol/2,
                     3*maxcol/4,
                     maxcol),
                   viridis(5)),
  na_col = 'white', 
  border = TRUE,
  border_gp = gpar(lwd = 0.1),
  column_names_gp = gpar(fontsize = 10),
  row_dend_width = unit(3, "cm"),
  width = unit(4.5, 'cm'),
  height = unit(7,'cm'),
  row_title = NULL,
  left_annotation = chapl_aapl_name_ha,
  row_split = 
    factor(combined_data_pir_ht$chapl_aapl_name,
              levels = ordering(combined_data_pir_ht$chapl_aapl_name,
                                order_vector = 'ChAPL')),
  heatmap_legend_param = list(
    labels_gp = gpar(fontsize = 10),
    title_gp = gpar(fontsize = 10,
                    fontface = 'bold'),
    at = c(0,
           maxcol/4,
           maxcol/2,
           3*maxcol/4,
           maxcol)
  )
)

chapl_aapl_name_lgd <- Legend(
  title = 'sub-family',
  at = factor(labels_chapl_aapl_name,
              ordering(labels_chapl_aapl_name,
                order_vector = 'ChAPL')),
  legend_gp = gpar(fill = annotation_col)
)

@@

```{r Supp_fig_AapL-ChaAPl_expression}

cir_plot <- twentyfour_rMT %>% 
  filter(Geneid %in% cir_id) %>%
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(
    stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
    diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,2],
    time = str_split_fixed(sample, pattern = '_', n = 3)[,1],
    subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)],
    chapl_aapl = factor(cir_info$chapl_aapl_name[match(Geneid, 
                                                       cir_info$Geneid)],
                        levels = sort(unique(cir_info$chapl_aapl_name[match(Geneid, 
                                                                            cir_info$Geneid)])))) %>% 
  filter(!(time %in% c('sch', 
                       'Unknown')),
         !is.na(chapl_aapl) ) %>% 
  group_by(Geneid) %>% 
    arrange(desc(tpm))

  cir_plot_order <- cir_plot %>% 
    filter(time == '14h') %>% 
    arrange(desc(tpm))

  cir_plot <- mutate(cir_plot, 
                     Geneid = factor(Geneid, 
                                     levels = unique(cir_plot_order$Geneid)))
  cir_plot_chapl_aapl_supp_info <- cir_plot

  max_aapl_chapl <- ceiling(max(log2(cir_plot$tpm+1)))

  breaks = seq(0, max_aapl_chapl, by = 1)
    limits = c(0, max_aapl_chapl)
    point_size = 0.25

  lapply(
    sort(unique(cir_plot$time)),
    function(time_vec){

    cir_plot <- filter(cir_plot, 
                       time == time_vec)

      cir_plot$tpm <- log2(1+cir_plot$tpm)

    cir_plot$Geneid <- factor(
      str_remove(cir_plot$Geneid, 
                 pattern = 'PCHAS_'),
      levels = str_remove(levels(cir_plot$Geneid), 
                          pattern = 'PCHAS_')
    )

    y <- ggplot(
      data = cir_plot,
      mapping = aes(
        x = Geneid,
        y = tpm
      )
    ) +
      stat_summary(fun.y = median, 
                   fun.ymin = median, 
                   fun.ymax = median,
                   geom = "bar", 
                   width = 0.75, 
                   size = 0.25, 
                   position = 'dodge',
                   fill = 'lightblue') +
      geom_point(stat = 'identity', 
                 position = position_dodge(width = 0.75),
                 size = point_size) +
      scale_y_continuous(breaks = breaks,
                         limits = limits) +
      scale_fill_manual(values = c('skyblue','red')) +
      ylab('log(TPM +1)') +
      facet_grid(.~chapl_aapl, 
                 scales = "free_x",
                 space = "free_x") +
      theme_classic() +
      theme(
        axis.text.x = element_text(angle = 90,
                                   size = 5)
      ) +
      ggtitle(paste(unique(cir_plot$time)))

    if(global_save_var){
      ggsave2(
        paste0('Pchabaudi_sibe-by-side_chapl_aapl_',
               time_vec,
               '.png'),
        plot = y,
        dpi = 300,
        width = 25,
        height = 10,
        units = 'cm'
      )
    }

    return(y)
    })
```

>>>>

combined_data_ordered <- tibble(Geneid = rownames(combined_data_pir),
                                means = apply(combined_data_pir[,
                                                                str_detect(colnames(combined_data_pir), 
                                                                           'chapl_aapl',
                                                                           negate = TRUE)],
                                              1,
                                              mean),
                                chapl_aapl_name = factor(
                                  combined_data_pir$chapl_aapl_name
                                )) %>% 
  arrange(chapl_aapl_name, 
          desc(means)) %>% 
  .$Geneid

@@

#Sub-family annotation
chapl_aapl_name_info <- combined_data_pir_ht$chapl_aapl_name
annotation_col <- chapl_aapl_name_col_num[ordering(labels_chapl_aapl_name,
                                                    order_vector = 'ChAPL')]

@@

ht <- Heatmap(
  log2(1+combined_data_pir_ht[,!colnames(combined_data_pir_ht) %in% 'chapl_aapl_name']),
  cluster_rows = FALSE,
  cluster_columns = FALSE,
  show_row_names = FALSE,
  name = 'log(TPM+1)', 
  col = colorRamp2(c(0,
                     maxcol/4,
                     maxcol/2,
                     3*maxcol/4,
                     maxcol),
                   viridis(5)),
  na_col = 'white', 
  border = TRUE,
  border_gp = gpar(lwd = 0.1),
  column_names_gp = gpar(fontsize = 10),
  row_dend_width = unit(3, "cm"),
  width = unit(4.5, 'cm'),
  height = unit(7,'cm'),
  row_title = NULL,
  left_annotation = chapl_aapl_name_ha,
  row_split =
    factor(combined_data_pir_ht$chapl_aapl_name,
              levels = ordering(combined_data_pir_ht$chapl_aapl_name,
                                order_vector = 'ChAPL')),
  heatmap_legend_param = list(
    labels_gp = gpar(fontsize = 10),
    title_gp = gpar(fontsize = 10,
                    fontface = 'bold'),
    at = c(0,
           maxcol/4,
           maxcol/2,
           3*maxcol/4,
           maxcol)
  ),
  row_gap = unit(c(rep.int(1, times = 2),
                   3,
                   rep.int(1, times = 4)), "mm")
)

chapl_aapl_name_lgd <- Legend(
  title = 'sub-family',
  at = str_subset(
    sort(
      factor(labels_chapl_aapl_name,
             ordering(labels_chapl_aapl_name,
                      order_vector = 'ChAPL'))
    ),
    pattern = 'NA',
    negate = TRUE
  ),
  legend_gp = gpar(fill = na.omit(annotation_col))
)

@@

```{r}
ht <- Heatmap(
  log2(1+combined_data_pir_ht[,!colnames(combined_data_pir_ht) %in% 'chapl_aapl_name']),
  cluster_rows = TRUE,
  cluster_columns = FALSE,
  show_row_names = FALSE,
  name = 'log(TPM+1)', 
  col = colorRamp2(c(0,
                     maxcol/4,
                     maxcol/2,
                     3*maxcol/4,
                     maxcol),
                   viridis(5)),
  na_col = 'white', 
  border = TRUE,
  border_gp = gpar(lwd = 0.1),
  column_names_gp = gpar(fontsize = 10),
  row_dend_width = unit(3, "cm"),
  width = unit(4.5, 'cm'),
  height = unit(7,'cm'),
  row_title = NULL,
  left_annotation = chapl_aapl_name_ha,
  # row_split = 
  #   factor(combined_data_pir_ht$chapl_aapl_name,
  #             levels = ordering(combined_data_pir_ht$chapl_aapl_name,
  #                               order_vector = 'ChAPL')),
  row_km = 9,
  heatmap_legend_param = list(
    labels_gp = gpar(fontsize = 10),
    title_gp = gpar(fontsize = 10,
                    fontface = 'bold'),
    at = c(0,
           maxcol/4,
           maxcol/2,
           3*maxcol/4,
           maxcol)
  )
)

chapl_aapl_name_lgd <- Legend(
  title = 'sub-family',
  at = str_subset(
    sort(
      factor(labels_chapl_aapl_name,
             ordering(labels_chapl_aapl_name,
                      order_vector = 'ChAPL'))
    ),
    pattern = 'NA',
    negate = TRUE
  ),
  legend_gp = gpar(fill = na.omit(annotation_col))
)
